### 2015-03-20 Kokkos_PG_HierarchicalParallelism.tex
Replaced [=] with KOKKOS_LAMBDA; explained Kokkos::single better.
I added an explanation of the KOKKOS_LAMBDA macro.  If CUDA is OFF, it
just turns into [=].  If CUDA is ON, it may do something special to make
the lambda work with CUDA.

I also improved the explanation of Kokkos::single.  In particular,
Kokkos::single's lambda is NOT allowed to modify any captured variable
by reference.

\subsection{Creating a Policy Instance}\label{SS:Hierarchical:Teams:Policy}

@@

\subsection{Basic Kernels}\label{SS:Hierarchical:Teams:Kernels}

Within the kernel the team policy member type provides the necessary functionality to use teams. 
It allows access to thread identifiers such as league rank and size and team rank and size, as well as team synchronous actions such as team barriers, reductions and scans.
\begin{lstlisting}
using Kokkos::TeamPolicy;
using Kokkos::parallel_for;

@@

This approach of writing hierarchical parallelism algorithms makes it very explicit that a kernel using a \lstinline!TeamPolicy! constitutes a parallel region with respect to the team.

@@

The most simple use case is to have another \lstinline|parallel_for| nested inside a kernel. 
\begin{lstlisting}
Kokkos::parallel_for(Kokkos::TeamPolicy<>(league_size,team_size), 
                     KOKKOS_LAMBDA ( member_type team_member) {
  Scalar tmp;
  Kokkos::parallel_for(
    Kokkos::TeamThreadRange(team_member,loop_count), 
    [=] (int& i) {
    ....
    // tmp += i; // This would be an illegal access
  });
});
\end{lstlisting}

@@

\subsection{Execution Restriction with \lstinline|single|}

As stated in the earlier subsections, a kernel has to be treated as a parallel region with respect to threads (and vector lanes) within a team.
This means that global memory accesses outside of the respective nested levels potentially have to be protected against repetitive execution. 
A common example is the case where a team performs some calculation but only one result per team has to be written back to global memory. 
Kokkos provides the \lstinline|Kokkkos::single(Policy,Lambda)| function for this case.
The two policies \lstinline|Kokkos::PerTeam| and \lstinline|Kokkos::PerThread| restrict execution to once per thread or once per team.

\begin{lstlisting}
using Kokkos::parallel_for;
using Kokkos::parallel_reduce;
using Kokkos::TeamThreadRange;
using Kokkos::ThreadVectorRange;
using Kokkos::PerThread;
// ...
parallel_for (TeamThreadRange (thread, 100), [=] (int i) {
    double sum = 0;
    // Perform a vector reduction with a thread
    parallel_reduce (ThreadVectorRange (thread, 100), 
      [=] (int i, double& lsum) {
        // ...
        lsum += ...;
      }, sum);
    // Add the result value into a team shared array.
    // Make sure it is only added once per thread.
    Kokkos::single (PerThread (), [=] () {
        shared_array(i) += sum;
      });
  });

double sum;
parallel_reduce (TeamThreadRange (thread, 99), 
  [=] (int i, double& lsum) {
    // Add the result value into a team shared array
    // Make sure its only added once per thread
    Kokkos::single (PerThread (thread), [=] () {
        lsum += someFunction (shared_array(i), 
                              shared_array(i+1));
      });
  }, sum);

// Add the per team contribution to global memory
Kokkos::single (PerTeam (thread), [=] () {
    global_array(thread.league_rank()) = sum;
  });

// ...
\end{lstlisting}

>>>>

\subsection{Creating a Policy instance}\label{SS:Hierarchical:Teams:Policy}

@@

\subsection{Basic kernels}\label{SS:Hierarchical:Teams:Kernels}

The team policy's \lstinline!member_type! provides the necessary functionality to use teams within a parallel kernel.
It allows access to thread identifiers such as the league rank and size, and the team rank and size.
It also provides team-synchronous actions such as team barriers, reductions and scans.
\begin{lstlisting}
using Kokkos::TeamPolicy;
using Kokkos::parallel_for;

@@

The name ``\lstinline!TeamPolicy!'' makes it explicit that a kernel
using it constitutes a parallel region with respect to the team.

@@

The simplest use case is to have another \lstinline|parallel_for| nested inside a kernel. 
\begin{lstlisting}
using Kokkos::parallel_for;
using Kokkos::TeamPolicy;
using Kokkos::TeamThreadRange;

parallel_for (TeamPolicy<> (league_size, team_size), 
                    KOKKOS_LAMBDA (member_type team_member)
{
  Scalar tmp;
  parallel_for (TeamThreadRange (team_member, loop_count), 
    [=] (int& i) {
      // ...
      // tmp += i; // This would be an illegal access
    });
});
\end{lstlisting}

@@

\subsection{Restrict execution with \lstinline|single|}

As stated above, a kernel is a parallel region with respect to threads (and vector lanes) within a team.
This means that global memory accesses outside of the respective nested levels potentially have to be protected against repetitive execution. 
A common example is the case where a team performs some calculation but only one result per team has to be written back to global memory.

Kokkos provides the \lstinline|Kokkkos::single(Policy,Lambda)| function for this case.
It currently accepts two policies:
\begin{itemize}
\item \lstinline|Kokkos::PerTeam| restricts execution of the lambda's
  body to once per team
\item \lstinline|Kokkos::PerThread| restricts execution of the
  lambda's body to once per thread (that is, to only one vector lane
  in a thread)
\end{itemize}
The \lstinline!single! function takes a lambda as its second argument.
That lambda takes zero arguments, 
so its body must perform side effects in order to have an effect.
It must always be correct for the lambda to capture variables by value
(\lstinline![=]!, not \lstinline![&]!).
Thus, if the lambda captures by reference,
it must \emph{not} modify variables that it has captured by reference.

\begin{lstlisting}
using Kokkos::parallel_for;
using Kokkos::parallel_reduce;
using Kokkos::TeamThreadRange;
using Kokkos::ThreadVectorRange;
using Kokkos::PerThread;

TeamPolicy<...> policy (...);
typedef TeamPolicy<...>::member_type team_member;

parallel_for (policy, KOKKOS_LAMBDA (const team_member& thread) {
  // ...

  parallel_for (TeamThreadRange (thread, 100), 
    KOKKOS_LAMBDA (const int& i) {
      double sum = 0;
      // Perform a vector reduction with a thread
      parallel_reduce (ThreadVectorRange (thread, 100), 
        [=] (int i, double& lsum) {
          // ...
          lsum += ...;
        }, sum);
      // Add the result value into a team shared array.
      // Make sure it is only added once per thread.
      Kokkos::single (PerThread (), [=] () {
          shared_array(i) += sum;
        });
    });

  double sum;
  parallel_reduce (TeamThreadRange (thread, 99), 
    KOKKOS_LAMBDA (int i, double& lsum) {
      // Add the result value into a team shared array.
      // Make sure it is only added once per thread.
      Kokkos::single (PerThread (thread), [=] () {
          lsum += someFunction (shared_array(i), 
                                shared_array(i+1));
        });
    }, sum);

  // Add the per team contribution to global memory.
  Kokkos::single (PerTeam (thread), [=] () {
      global_array(thread.league_rank()) = sum;
    });
  });
\end{lstlisting}

### 2015-03-20 Kokkos_PG_ParallelDispatch.tex
Minor changes to section labels.
Replaced [=] with KOKKOS_LAMBDA; explained Kokkos::single better.
I added an explanation of the KOKKOS_LAMBDA macro.  If CUDA is OFF, it
just turns into [=].  If CUDA is ON, it may do something special to make
the lambda work with CUDA.

I also improved the explanation of Kokkos::single.  In particular,
Kokkos::single's lambda is NOT allowed to modify any captured variable
by reference.

\chapter{Parallel dispatch}

You probably started reading this Guide because you wanted to learn
how Kokkos can parallelize your code.  This chapter will teach you
different kinds of parallel operations that Kokkos can execute.  We
call these operations collectively \emph{parallel dispatch}, because
Kokkos ``dispatches'' them for execution by a particular execution
space.  Kokkos provides three different parallel operations:
\begin{itemize}
\item \lstinline!parallel_for! implements a ``for loop'' with
  independent iterations.
\item \lstinline!parallel_reduce! implements a reduction.
\item \lstinline!parallel_scan! implements a prefix scan.
\end{itemize}
Kokkos gives users two options for defining the body of a parallel
loop: functors and lambdas.  It also lets users control how the
parallel operation executes, by specifying an \emph{execution policy}.

@@

By default, lambdas capture nothing (as the default capture specifier \lstinline![]! specifies).
This is not likely to be useful, since parallel for generally works by its side effects.
Thus, we recommend using the ``capture by value'' specifier \lstinline![=]! by default.
You may also explicitly specify variables to capture, but they must be captured by value.

@@

Here is an example reduction using a lambda,
where the reduction result is a \lstinline!double!.
\begin{lstlisting}
const size_t N = ...;
View<double*> x ("x", N);
// ... fill x with some numbers ... 
double sum = 0.0;
parallel_reduce (N, [=] (const int i, double& update) {
    update += x(i); }, sum);
\end{lstlisting}
This version of \verb!parallel_reduce! is easy to use,
but it imposes some assumptions on the reduction.
For example, it assumes that it is correct for threads to join their intermediate reduction results using binary \lstinline!operator+!.
If you want to change this, you must either implement your own reduction result type with a custom binary \lstinline!operator+!,
or define the reduction using a functor instead of a lambda.

@@

// Use the above functor to compute the sum of squares from 0 to N-1.
const size_t N = ...;
sum = 0;
// parallel_reduce needs an instance of SquareSum, 
// so we invoke its constructor with ().
parallel_reduce (N, SquareSum ());
\end{lstlisting}
This example has a short enough loop body that it would be better to use a lambda:
\begin{lstlisting}
// Compute the sum of squares from 0 to N-1.
int sum = 0;
parallel_reduce (N, [=] (const int i, int& lsum) { lsum += i*i; });
\end{lstlisting}

@@

Kokkos lets users specify a scan by either a functor or a lambda.
Both look like their \lstinline!parallel_reduce! equivalents,
except that the \lstinline!operator()! method or lambda takes three arguments:
the loop index, the ``update'' value by nonconst reference, and a \lstinline!bool!.
Here is a lambda example where the intermediate results have type \lstinline!float!.
\begin{lstlisting}
View<float*> x = ...; // assume filled with input values
const size_t N = x.dimension_0 ();
parallel_scan (N, [=] (const int i, float& upd, const bool final) {
    if (final) {
      x(i) = upd; // only update array on final pass
    }
    // For exclusive scan, change the update value after 
    // updating array, like we do here.  For inclusive scan, 
    // change the update value before updating array.
    upd += x(i); 
  });
\end{lstlisting}
Kokkos may use a multiple-pass algorithm to implement scan.
This means that it may call your \lstinline!operator()! or lambda multiple times per loop index value.
The \lstinline!final! Boolean argument tells you whether Kokkos is on the final pass.
You must only update the array on the final pass.

>>>>

\chapter{Parallel dispatch}\label{C:Dispatch}

You probably started reading this Guide because you wanted to learn
how Kokkos can parallelize your code.  This chapter will teach you
different kinds of parallel operations that Kokkos can execute.  We
call these operations collectively \emph{parallel dispatch}, because
Kokkos ``dispatches'' them for execution by a particular execution
space.  Kokkos provides three different parallel operations:
\begin{itemize}
\item \lstinline!parallel_for! implements a ``for loop'' with
  independent iterations.
\item \lstinline!parallel_reduce! implements a reduction.
\item \lstinline!parallel_scan! implements a prefix scan.
\end{itemize}
Kokkos gives users two options for defining the body of a parallel
loop: functors and lambdas.  It also lets users control how the
parallel operation executes, by specifying an \emph{execution policy}.
Later chapters will cover more advanced execution policies that allow
nested parallelism.

Important notes on syntax:
\begin{itemize}
\item Use the \lstinline!KOKKOS_INLINE_FUNCTION! macro to mark a
  functor's methods that Kokkos will call in parallel
\item Use the \lstinline!KOKKOS_LAMBDA! macro to replace a lambda's
  capture clause when giving the lambda to Kokkos for parallel
  execution
\end{itemize}

@@

By default, lambdas capture nothing (as the default capture specifier \lstinline![]! specifies).
This is not likely to be useful, since parallel for generally works by its side effects.
Thus, we recommend using the ``capture by value'' specifier \lstinline![=]! by default.
You may also explicitly specify variables to capture, but they must be captured by value.
We prefer that for the outermost level of parallelism (see Chapter \ref{C:Hierarchical}),
you use the \lstinline!KOKKOS_LAMBDA! macro instead of the capture clause.
If CUDA is disabled, this just turns into the usual capture-by-value clause \lstinline![=]!.
That captures variables from the surrounding scope by value.
Do NOT capture them by reference!
If CUDA is enabled, this macro may have a special definition 
that makes the lambda work correctly with CUDA.
Compare to the \lstinline!KOKKOS_INLINE_FUNCTION! macro,
which has a special meaning if CUDA is enabled.
If you do not plan to build with CUDA, you may use \lstinline![=]! explicitly,
but we find using the macro easier than remembering the capture clause syntax.

@@

Here is an example reduction using a lambda,
where the reduction result is a \lstinline!double!.
\begin{lstlisting}
const size_t N = ...;
View<double*> x ("x", N);
// ... fill x with some numbers ... 
double sum = 0.0;
// KOKKOS_LAMBDA macro includes capture-by-value specifier [=].
parallel_reduce (N, KOKKOS_LAMBDA (const int i, double& update) {
    update += x(i); }, sum);
\end{lstlisting}
This version of \verb!parallel_reduce! is easy to use,
but it imposes some assumptions on the reduction.
For example, it assumes that it is correct for threads to join their intermediate reduction results using binary \lstinline!operator+!.
If you want to change this, you must either implement your own reduction result type with a custom binary \lstinline!operator+!,
or define the reduction using a functor instead of a lambda.

@@

// Use the above functor to compute the sum of squares from 0 to N-1.
const size_t N = ...;
sum = 0;
// parallel_reduce needs an instance of SquareSum, 
// so we invoke its constructor with ().
parallel_reduce (N, SquareSum ());
\end{lstlisting}
This example has a short enough loop body that it would be better to use a lambda:
\begin{lstlisting}
// Compute the sum of squares from 0 to N-1.
int sum = 0;
parallel_reduce (N, KOKKOS_LAMBDA (const int i, int& lsum) { lsum += i*i; });
\end{lstlisting}

@@

Kokkos lets users specify a scan by either a functor or a lambda.
Both look like their \lstinline!parallel_reduce! equivalents,
except that the \lstinline!operator()! method or lambda takes three arguments:
the loop index, the ``update'' value by nonconst reference, and a \lstinline!bool!.
Here is a lambda example where the intermediate results have type \lstinline!float!.
\begin{lstlisting}
View<float*> x = ...; // assume filled with input values
const size_t N = x.dimension_0 ();
parallel_scan (N, KOKKOS_LAMBDA (const int i, float& upd, const bool final) {
    if (final) {
      x(i) = upd; // only update array on final pass
    }
    // For exclusive scan, change the update value after 
    // updating array, like we do here.  For inclusive scan, 
    // change the update value before updating array.
    upd += x(i); 
  });
\end{lstlisting}
Kokkos may use a multiple-pass algorithm to implement scan.
This means that it may call your \lstinline!operator()! or lambda multiple times per loop index value.
The \lstinline!final! Boolean argument tells you whether Kokkos is on the final pass.
You must only update the array on the final pass.

### 2015-03-20 Kokkos_PG_Views.tex
Replaced [=] with KOKKOS_LAMBDA; explained Kokkos::single better.
I added an explanation of the KOKKOS_LAMBDA macro.  If CUDA is OFF, it
just turns into [=].  If CUDA is ON, it may do something special to make
the lambda work with CUDA.

I also improved the explanation of Kokkos::single.  In particular,
Kokkos::single's lambda is NOT allowed to modify any captured variable
by reference.

You may access an entry of a View using parentheses enclosing a
comma-delimited list of integer indices.  This looks just like a
Fortran multidimensional array access.  For example:
\begin{lstlisting}
const size_t N = ...;
View<double*[3][4]> a ("some label", N);
parallel_for (N, [=] (const ptrdiff_t i) {
    const size_t j = ...;
    const size_t k = ...;
    const double a_ijk = a(i,j,k);
    /* rest of the loop body */
  });
\end{lstlisting}
Note how in the above example, we only access the View's entries in a
parallel loop body.  In general, you may only access a View's entries
in an execution space which is allowed to access that View's memory
space.  For example, if the default memory space is \lstinline!Cuda!
(CUDA device memory), you may not access it in host code, unless CUDA
UVM is enabled.  Furthermore, access costs (e.g., latency and
bandwidth) may vary, depending on the View's ``native'' memory and
execution spaces, and the execution space from which you access it.
CUDA UVM may work, but it may also be slow, depending on your access
pattern and performance requirements.  Thus, best practice is to
access the View only in a Kokkos parallel for, reduce, or scan, using
the same execution space as the View.  This also ensures that access
to the View's entries respect first-touch allocation.  The first
(leftmost) dimension of the View is the \emph{parallel dimension},
over which it is most efficient to do parallel array access.

@@

  parallel_for (y.dimension_0 (), [=] (const size_type i) {
      double y_i = y(i);
      for (size_t k = ptr(i); k < ptr(i+1); ++k) {
        y_i += val(k) * x_ra(ind(k));
      }
      y(i) = y_i;
    });
}
\end{lstlisting}

>>>>

You may access an entry of a View using parentheses enclosing a
comma-delimited list of integer indices.  This looks just like a
Fortran multidimensional array access.  For example:
\begin{lstlisting}
const size_t N = ...;
View<double*[3][4]> a ("some label", N);
// KOKKOS_LAMBDA macro includes capture-by-value specifier [=].
parallel_for (N, KOKKOS_LAMBDA (const ptrdiff_t i) {
    const size_t j = ...;
    const size_t k = ...;
    const double a_ijk = a(i,j,k);
    /* rest of the loop body */
  });
\end{lstlisting}
Note how in the above example, we only access the View's entries in a
parallel loop body.  In general, you may only access a View's entries
in an execution space which is allowed to access that View's memory
space.  For example, if the default memory space is \lstinline!Cuda!
(CUDA device memory), you may not access it in host code, unless CUDA
UVM is enabled.  Furthermore, access costs (e.g., latency and
bandwidth) may vary, depending on the View's ``native'' memory and
execution spaces, and the execution space from which you access it.
CUDA UVM may work, but it may also be slow, depending on your access
pattern and performance requirements.  Thus, best practice is to
access the View only in a Kokkos parallel for, reduce, or scan, using
the same execution space as the View.  This also ensures that access
to the View's entries respect first-touch allocation.  The first
(leftmost) dimension of the View is the \emph{parallel dimension},
over which it is most efficient to do parallel array access.

@@

  parallel_for (y.dimension_0 (), KOKKOS_LAMBDA (const size_type i) {
      double y_i = y(i);
      for (size_t k = ptr(i); k < ptr(i+1); ++k) {
        y_i += val(k) * x_ra(ind(k));
      }
      y(i) = y_i;
    });
}
\end{lstlisting}
